Empirical Bayes analysis of \RNAseq{} data for detection of gene expression heterosis

### 2015-09-01 response.tex

\comment{The comments made by one reviewer regarding a relative assessment of computing time relative to the other methods deserves attention.}

\response{Added. See section 4.}

\comment{Page 1.  I was instructed that heterosis is merely characterized by the hybrid performance not being equal to the average of the two parental lines.  So in your very first sentence, I might write “Heterosis, or hybrid vigor, occurs when hybrid progeny display phenotypes that are superior to the average phenotypes of their parents.” i.e., I might be only interested in the $\delta_g$’s. However, your later characterization of heterosis is really something called overdominance to me…where the hybrid performance exceeds the performance of either parent.  I wonder if you could then be more specific in your definitions in your first sentence…and also in your third sentence in this particular paragraph. lines 48-59.  Perhaps this was already captured well in the Ji et al. (2014) paper but why not simply test heterosis as the difference between the hybrid and the average parental performance (i.e. $\mu_3 - 0.5 \mu_1 - 0.5 \mu_2$).  Again, I think your issue pertains really to an overdominance test of some sort.}

\response{Many publications mentioning heterosis do not define it clearly.  For example, the 2013 {\em Nature Reviews Genetics} paper by Chen that we now cite in our revised manuscript states the following definition: ``Heterosis (Also known as hybrid vigour).  When hybrids display increased levels of growth, survival or fitness relative to their parents.''  Like Chen, some other author present heterosis as synonymous with hybrid vigor, and then go on to describe hybrid vigor as the phenomenon in which the hybrid outperforms the parents.  It may not be so clear what ``outperforms the parents'' means, but the idea behind hybrid vigor and the reason it is of such interest (particularly in maize) and so useful is that the hybrid is better than {\bf both parents}, not simply better than the average.  That said, we do agree with your definition of heterosis.  We originally wrote that first sentence so that it would be true regardless of which definition of heterosis one takes.  We were saying that heterosis occurs when hybrid vigor occurs.  The reverse is not necessarily so, and we did not claim it to be in that first sentence.  However, we acknowledge that our first sentence was misleading because most readers would naturally take the sentence for a definition.  Furthermore, for simplicity, we did start using ``heterosis'' as a synonym for the situation where the hybrid mean expression is more extreme than both parents.  That is really what we are interested in because the phenotypic traits we see in maize have this pattern, so it is only natural to seek genes whose expression has this same pattern.  In the revised manuscript, we have been more careful with terminology.  We have defined heterosis to be a difference between the hybrid mean the average of the parental means.  We have also introduced a new term (extreme heterosis, EH) to specify the situation where the hybrid mean is more extreme than both parental means.  We continue to use low-parent heterosis (LPH) and high-parent heterosis (HPH) to distinguish the two types of EH.  We also explain in the manuscript why we prefer the LPH/HPH terminology to underdominance/overdominance.  In our experience, these latter terms are used to describe the ``gene action'' at a single locus with respect to some quantitative trait.  We can't asses the action of any locus on any of our expression traits in data from inbred and hybrid crosses.}

### 2015-09-01 ebayes_rnaseq_heterosis.tex
last paragraph

Notwithstanding these improvements, we believe our approach is a computationally efficient method that can immediately aid scientists who are interested in identifying candidate genes for a genetic mechanism of heterosis.

An alternative explanation for phenotypic heterosis is preferential gene expression of alleles from certain parents for certain genes. Analyses of allele-specific expression \citep{bell2013rna,wei2013computational} could complement the search for gene expression heterosis in the quest for understanding the molecular genetic mechanisms behind phenotypic heterosis.

>>>>

Notwithstanding these improvements, we believe our approach is a computationally efficient method that can immediately aid scientists who are interested in identifying candidate genes involved in a genetic mechanism of heterosis.

This paper has focused on statistical methods for detecting EH in gene expression using RNA-seq data.  EH at the transcript level is only one of multiple molecular genetic mechanisms that may play roles in establishing hybrid vigor.  Complementation \citep{paschold2012complementation}, allele-specific expression \citep{bell2013rna,wei2013computational}, and other complex forms of genomic and epigenetic interaction \citep{chen2013genomic} are all plausible as mechanisms partially responsible for phenotypic heterosis.  Further study of these phenomena using modern genomic technologies and appropriate statistical methods should enhance our understanding of hybrid vigor.
